Estimating duration using survival analysis

### 2024-03-29 cis-perfect-testing.tex
Updates from Overleaf

In this section, I develop a computationally feasible Bayesian statistical model accounting for double interval censoring and undetected episodes.
In \cref{perf-test:sec:data}, I introduce the data, defining which quantities are observed and which are latent.
Then, in \cref{perf-test:sec:episode-types}, I build a framework, using the concept of \textcite{heiseyModelling}, which classifies possible episode beginning and end time into \emph{admissible} regions, each of which match a possible pattern of the doubly interval censored pattern of the observations; and an \emph{undetected} region, which lead to the episode being undetected.
The number of such episodes in each region has a multinomial distribution conditional on the latent total number of infections (including undetected infections).

@@

I analyse the $\ndet = \numprint{4800}$ episodes with their first positive test between 10 Oct 2020 and 6 Dec 2020 inclusive, with negatives bounding the start and end time of the episode; I refer to these episodes as \emph{detected}.
In \cref{perf-test:sec:prob-undetected} I formalise the conditions for an episode to be detected.
This period is denoted as $t = 1, \dots, T$.
A negative test before the start of the episode provides a lower bound on the episode's start time; otherwise, there is little information on its length.
Similar information comes from the negative test after the end of the episode.
A negligible number of episodes are missing a negative test after the end of the episode.

@@

@@

Any infection episode, whether detected or not, is fully characterised by the triplet $(b_j, e_j, i(j))$ where $b_j$ and $e_j$ are the times the episode begin and end respectively; as $i(j)$ is the individual the episode occurred in it also identifies the test schedule.
The episode's duration is $d_j = e_j - b_j + 1$ (see \cref{E-inc-prev:sec:definitions}).
For all $j$, the exact values $b_j$ and $e_j$ are latent.
If $j$ is undetected then $i(j)$ is also latent and little is known about $b_j$ or $e_j$.
If $j$ is detected, then $i(j)$ is known; $b_j \in [l_j^{(b)}, r_j^{(b)}]$; and $e_j \in [l_j^{(e)}, r_j^{(e)}]$.
These latter two properties are what make the data doubly interval censored.

@@

In the CIS data, each $n_k$ ($k \neq u$) is observed as either 0 or 1.
Define $D = \{ k \ssep n_k = 1 \}$.
Furthermore, note that the support of the multinomial distribution requires that $\nnodet = \ntot - \ndet$.
Let $\na = [n_{1}, \dots, n_{N_E}]^T$ be the observed portion of $\vec{n}$.
Then \cref{perf-test:eq:multinomial-ll} simplifies to:
\begin{align}
  p(\vec{n} \mid \ntot, \vec{\theta})
  &= p(\na \mid \ntot, \vec{\theta}) \\
  &= \frac{\ntot!}{(\ntot - \ndet)!} p_u^{\ntot-\ndet} \prod_{k \in \set{D}} p_k.
  \label{perf-test:eq:multinomial}
\end{align}

>>>>

In this section, I develop a computationally feasible Bayesian statistical model accounting for double interval censoring and undetected episodes.
In \cref{perf-test:sec:data}, I introduce the data, defining which quantities are observed and which are latent.
Then, in \cref{perf-test:sec:posterior}, I build a framework, using the concept of \textcite{heiseyModelling}.
I classify episode based on the possible patterns of the doubly interval censored pattern of the observations; and an \emph{undetected} region, which lead to the episode being undetected.
The number of such episodes in each region has a multinomial distribution conditional on the latent total number of infections (including undetected infections).

@@

I analyse the $\ndet = \numprint{4800}$ episodes with their first positive test between 10 Oct 2020 and 6 Dec 2020 inclusive, with negatives bounding the start and end time of the episode; I refer to these episodes as \emph{detected}.
In \cref{perf-test:sec:prob-undetected} I formalise the conditions for an episode to be detected.
Each day in this period is denoted as $t = 1, \dots, T$.
A negative test before the start of the episode provides a lower bound on the episode's start time; otherwise, there is little information on its length.
Similar information comes from the negative test after the end of the episode.
A negligible number of episodes are missing a negative test after the end of the episode.

@@

Any infection episode, whether detected or not, is fully characterised by the triplet $(b_j, e_j, i(j))$ where $b_j$ and $e_j$ are the times the episode begin and end respectively; and $i(j)$ is the individual the episode occurred in.
$i(j)$ identifies the test schedule.
The episode's duration is $d_j = e_j - b_j + 1$ (see \cref{E-inc-prev:sec:definitions}).
For all $j$, the exact values $b_j$ and $e_j$ are latent.
If $j$ is undetected then $i(j)$ is also latent and little is known about $b_j$ or $e_j$.
If $j$ is detected, then $i(j)$ is known; $b_j \in [l_j^{(b)}, r_j^{(b)}]$; and $e_j \in [l_j^{(e)}, r_j^{(e)}]$.
These latter two properties are what make the data doubly interval censored.

@@

@@

In the CIS data, each $n_k$ ($k \neq u$) is observed as either 0 or 1.
Define $\set{D} = \{ k \ssep n_k = 1 \}$, the detected episodes.
Furthermore, note that the support of the multinomial distribution requires that $\nnodet = \ntot - \ndet$.
Let $\na = [n_{1}, \dots, n_{N_E}]^T$ be the observed portion of $\vec{n}$.
Then \cref{perf-test:eq:multinomial-ll} simplifies to:
\begin{align}
  p(\vec{n} \mid \ntot, \vec{\theta})
  &= p(\na \mid \ntot, \vec{\theta}) \\
  &= \frac{\ntot!}{(\ntot - \ndet)!} p_u^{\ntot-\ndet} \prod_{k \in \set{D}} p_k.
  \label{perf-test:eq:multinomial}
\end{align}
